---
title: "Homework 3"
author: "Jordan Atkins"
date: "2025-02-11"
output: 
 html_document: 
  toc: true
  toc_depth: 2
  toc_float: true
---

# Problem 1

## Theory A:

### Claim: Gas stations charge more if they lack direct competition in sight.

```{r echo=FALSE}
gas <- read.csv('gasprices.csv')

options(scipen=999)

library(tidyverse)

library(ggplot2)
library(dplyr) 

gas.gg <- ggplot(gas)

gas.gg +  geom_histogram(aes(x=Price), col='black', fill='green', bins=9) + 
  labs(title='Gas prices compared by proximity to competitors', x='Price ($/gal)', y='Frequency') + 
  theme(legend.position='bottom') + 
  scale_fill_brewer(type='qual', palette=2) +facet_wrap(~Competitors)

median(gas$Price[gas$Competitors == 'N'])
median(gas$Price[gas$Competitors == 'Y'])
```

### 

I used the histograms of the gas prices of stations near competition and those without as well as the medians of both of these graphs. The histogram for station with competition (y) is approximately normal and the stations without (n) are bi modal with more stations having prices near or above \$2. When comparing the medians of these graphs the median for stations with competition (y) was 1.85 and for those without was 1.89. Based on the evidence above I conclude that there is not significant evidence to support gas stations charging more when there are no competitors. Thes values are not significant enough to quantify a conclusion

## Theory B:

```{r echo=FALSE}

ggplot(gas, aes(x = Price, y = Income)) +
  geom_point() +
  labs(title = "Gase Prices by income",
       x = "Price(4)",
       y = "Average Income",
       color = "month") +
  theme_minimal() + scale_y_continuous() 

cor(gas$Price, gas$Income)

lm_gasincome<- lm(Price ~ Income, data = gas)
summary(lm_gasincome)
confint(lm_gasincome)
```

The scatter plot for gas prices compared by income has no linear trend and a correlation coefficient of 0.39 it seems that that there is no relationship. But a linear regression would produces significat p value. So I conclude with 95% confidence, the difference between gas stations at different incomes is between 0.0671 and 0.1826 cents.

## Theory C

### Gas stations at stoplights charge more

```{r echo=FALSE}
gStoplight <- as.factor(gas$Stoplight)

# Compute summary statistics
summary_by_Spotlight <- gas %>%
  group_by(Stoplight) %>%
  summarise(
    median_price = median(Price, na.rm = TRUE),
    sd_price = sd(Price, na.rm = TRUE),
    n = n()
  )

# Print summary
print(summary_by_Spotlight)

# Perform t-test to compare means
t_test_light <- t.test(Price ~ Stoplight, data = gas)
print(t_test_light)

# Extract confidence interval
ci_light<- t_test_light$conf.int
print(ci_light)

# Scatter plot with highway vs non-highway prices
ggplot(gas, aes(x = Stoplight, y = Price)) +
  geom_boxplot() +
  labs(title = "Gas Prices at Stoplights vs  not at stoplights",
       x = "Y = At a stoplight N = Not at Spotlight",
       y = "Gas Price ($)")
```

The evidence does not support the claim that gas stations at stoplights charge more. The observed difference in gas prices is almost zero (-0.3 cents), and the p-value is very high (0.8521), meaning the difference is likely due to chance. The 95% confidence interval includes zero, further confirming that there is no meaningful difference.

## Theory D

```{r echo=FALSE}
gas$Highway <- as.factor(gas$Highway)

# Compute summary statistics
summary_by_highway <- gas %>%
  group_by(Highway) %>%
  summarise(
    mean_price = mean(Price, na.rm = TRUE),
    sd_price = sd(Price, na.rm = TRUE),
    n = n()
  )

# Print summary
print(summary_by_highway)

# Perform t-test to compare means
t_test_highway <- t.test(Price ~ Highway, data = gas)
print(t_test_highway)

# Extract confidence interval
ci_highway <- t_test_highway$conf.int
print(ci_highway)

# Scatter plot with highway vs non-highway prices
ggplot(gas, aes(x = Highway, y = Price)) +
  geom_boxplot() +
  labs(title = "Gas Prices: Highway vs. Non-Highway",
       x = "Highway Access",
       y = "Gas Price ($)")

```

The boxplot for Highway vs non-highway stations hows stations with highway access to have slightly higher median gas price and high upper bounds and lower pounds he evidence suggests that gas stations on the highway charge significantly more (between 0.8 and 8.3 cents per gallon more, with 95% confidence). Since the p-value is below 0.05, we reject the null hypothesis and conclude that there is a significant price difference favoring higher prices on the highway.

## Theory E

### Shell charges more than all other non-Shell brand

```{r echo=FALSE, message=FALSE, warning=FALSE}
gas <- read_csv("gasprices.csv")

# Create a binary Shell vs. Non-Shell variable
gas$Shell <- ifelse(gas$Brand == "Shell", "Shell", "Other")
gas$Shell <- as.factor(gas$Shell)

# Compute summary statistics
summary_by_shell <- gas%>%
  group_by(Shell) %>%
  summarise(
    mean_price = mean(Price, na.rm = TRUE),
    sd_price = sd(Price, na.rm = TRUE),
    n = n()
  )


print(summary_by_shell)


t_test_shell <- t.test(Price ~ Shell, data = gas)
print(t_test_shell)


ci_shell <- t_test_shell$conf.int
print(ci_shell)

# Boxplot to visualize price difference
ggplot(gas, aes(x = Shell, y = Price)) +
  geom_boxplot() +
  labs(title = "Gas Prices: Shell vs. Other Brands",
       x = "Brand",
       y = "Gas Price ($)")
```

-   Mean gas price for non-Shell stations (Other) = \$1.85. Mean gas price for Shell stations = \$1.884. Difference in means = \$1.884 - \$1.856 = \$0.028 (2.8 cents). p-value = 0.1603, which is not statistically significant (p \> 0.05).95% Confidence Interval for the price difference = (-\$0.066, \$0.011). Since this confidence interval includes zero, it means there is no significant difference in gas prices between Shell and non-Shell brands. e evidence does not support the claim that Shell gas stations charge more than non-Shell brands. The observed difference in gas prices (2.8 cents higher at Shell) is not statistically significant (p = 0.1603, which is \> 0.05). Furthermore, the confidence interval includes zero, meaning that any difference could be due to random chance.

# Problem 2

## Part A

```{r echo=FALSE}
library(mosaic)
library(dplyr)
sclass <- read.csv('sclass.csv')

car <- sclass %>%
  filter(year == 2011, trim == "63 AMG")

bootstrap_samples <- do(10000) * mean(resample(car$mileage, replace = TRUE))

CI <- quantile(bootstrap_samples$mean, probs = c(0.025, 0.975))


print(CI)
```

With 95% confidence the average mileage for 2011 S-Class 63 AMG's is between 262945.73 and 31752.11\

## Part B

```{r echo = FALSE}

car2 <- mutate(sclass, isblack = color == "Black")
car3 <-  filter(car2, year == 2014 & trim == "550") 

boot_car<- do(10000) * prop(resample(car3$isblack, replace = TRUE))

# Compute the 95% Confidence Interval
car_ci<- quantile(boot_car$prop_TRUE, probs = c(0.025, 0.975), na.rm=TRUE)

print(car_ci)
                  
```

With 95% confidence i can predict that between 41% and 45% of 2014 S-Class 550's were painted Black

# Problem 3

## Part A

Between Living with ed and My name is Earl, does one show perform better with Happienes?

```{r echo=FALSE}
nbc <- read.csv('nbc_pilotsurvey.csv')

nbc1 <- nbc %>%
  filter(Show == "Living with Ed" | Show == "My Name is Earl")

nbc1_test <- t.test(Q1_Happy ~ Show, data = nbc1)

nbc1_test

```

Based on the t-test results, there is no significant difference in Q1_Happy scores between *Living with Ed* and *My Name is Earl*. The 95% confidence interval (-0.1008, 0.3989) includes 0, meaning the true difference in means could be positive or negative, making it statistically insignificant. Further analysis with a larger sample size or different response variables might be needed for stronger conclusions.

## Part B:  Annoyance Levels between The Biggest Loser” and “The Apprentice: Los Angeles.”

```{r echo=FALSE}

nbc2 <- nbc %>%
  filter(Show == "The Biggest Loser" | Show == "The Apprentice: Los Angeles")

nbc2_test <- t.test(Q1_Annoyed ~ Show, data = nbc2)

nbc2_test


```

The t-test results for Q1_Annoyed indicate a statistically significant difference in mean annoyance scores between *The Apprentice: Los Angeles* and *The Biggest Loser*. The mean annoyance score for *The Apprentice: Los Angeles* (2.31) is higher than that for *The Biggest Loser* (2.04), indicating that viewers found *The Apprentice: Los Angeles* significantly more annoying. Since the p-value is less than 0.05, we reject the null hypothesis and conclude that *The Apprentice: Los Angeles* tends to elicit higher annoyance ratings compared to *The Biggest Loser*.

## Part C: Do viewers Find Dancing with the stars Confusing?

```{r echo=FALSE}

nbc3 <- nbc %>%
  filter(Show == "Dancing with the Stars")
p_hat <- mean(nbc3$Q2_Confusing >= 4)

# Calculate the sample size
n <- nrow(nbc3)

# Compute the standard error (SE) for the proportion
SE <- sqrt((p_hat * (1 - p_hat)) / n)

# Compute the 95% confidence interval using the normal approximation (Wald method)
z_score <- 1.96  # Approximate Z-score for 95% confidence level
CI_lower <- p_hat - z_score * SE
CI_upper <- p_hat + z_score * SE

cat("Proportion of viewers rating Q2_Confusing ≥ 4:", round(p_hat, 4), "\n")
cat("95% Confidence Interval: (", round(CI_lower, 4), ",", round(CI_upper, 4), ")\n")

```

The analysis estimates that 7.73% of viewers rated *Dancing with the Stars* as confusing The 95% confidence interval for this proportion is (3.84%, 11.63%), meaning we are 95% confident that the true proportion of all American TV watchers who would give a confusion rating of 4 or higher falls within this range.

# Problem 4

```{r}

ebay <- read.csv("ebay.csv")

# Create a new column for revenue ratio
ebay <- ebay %>%
  mutate(revenue_ratio = rev_after / rev_before)

# Split data into treatment and control groups
treatment_group <- ebay %>% filter(adwords_pause == 1) %>% pull(revenue_ratio)
control_group <- ebay %>% filter(adwords_pause == 0) %>% pull(revenue_ratio)

# Compute observed difference in means
obs_diff <- mean(treatment_group) - mean(control_group)

# Bootstrap simulation (10,000 resamples)
set.seed(123)  # Set seed for reproducibility
n_simulations <- 10000
bootstrap_diffs <- replicate(n_simulations, {
  sample_treatment <- sample(treatment_group, size = length(treatment_group), replace = TRUE)
  sample_control <- sample(control_group, size = length(control_group), replace = TRUE)
  mean(sample_treatment) - mean(sample_control)
})

# Compute 95% confidence interval
CI <- quantile(bootstrap_diffs, probs = c(0.025, 0.975))

# Print results
cat("Observed Difference in Revenue Ratio:", obs_diff, "\n")
cat("95% Confidence Interval: (", CI[1], ",", CI[2], ")\n")

```

The observed difference in revenue ratio between the treatment group (AdWords paused) and the control group (AdWords active) is -0.0523. This means that, on average, the revenue ratio in the treatment group was 5.23 percentage points lower than in the control group. The results indicate that pausing paid search advertising on Google resulted in a significantly lower revenue ratio, suggesting that the ads were generating additional revenue for eBay. If paid search advertising had no effect, we would expect the revenue ratios to be similar between the treatment and control groups, but this is not the case. Instead, the confidence interval suggests that pausing ads led to a decline in revenue between 1.3% and 9.1%. Therefore, based on this analysis, paid search advertising on Google does contribute to extra revenue for eBay, and stopping it leads to a measurable decline in revenue performance.
